\textls[-0]{\thetitle}

### 2023-03-17 abstract.tex
tweaks

With the growing practice of mechanizing language metatheories,
it has become ever more pressing that interactive theorem provers 
make it easy to write reusable, extensible code and proofs.
%
This paper presents a novel language design geared towards extensible metatheory
mechanization in a proof assistant.
The new design achieves reuse and extensibility via a form of family
polymorphism, a seemingly object-oriented idea, that allows code and
proofs to be polymorphic to their enclosing families.
Our development addresses technical challenges that arise
from the underlying language of a proof assistant being simultaneously
functional, dependently typed, a logic, and an interactive tool.
%
Our results include
\begin{enumerate*}
\item a prototypical implementation of the language design as a Coq plugin,
\item a dependent type theory capturing the essence of the language mechanism
      and its consistency and canonicity results,
%\EDJ{We don't yet have normalization though}
and 
\item case studies showing how the new expressiveness naturally addresses real
programming challenges in metatheory mechanization.
\end{enumerate*}

>>>>

With the growing practice of mechanizing language metatheories,
it has become ever more pressing that interactive theorem provers 
make it easy to write reusable, extensible code and proofs.
%
This paper presents a novel language design geared towards extensible metatheory
mechanization in a proof assistant.
The new design achieves reuse and extensibility via a form of family
polymorphism, an object-oriented idea, that allows code and
proofs to be polymorphic to their enclosing families.
Our development addresses technical challenges that arise
from the underlying language of a proof assistant being simultaneously
functional, dependently typed, a logic, and an interactive tool.
%
Our results include
\begin{enumerate*}
\item a prototypical implementation of the language design as a Coq plugin,
\item a dependent type theory capturing the essence of the language mechanism
      and its consistency and canonicity results,
%\EDJ{We don't yet have normalization though}
and 
\item case studies showing how the new expressiveness naturally addresses real
programming challenges in metatheory mechanization.
\end{enumerate*}

### 2023-03-17 metatheory-condensed.tex
tweaks

Our formulation of W-types differs from previous ones in that it makes
individual constructors explicit and thus more closely matches the surface
syntax of \Lang.
%
In \TT, a W-type is given by a \emph{W-type signature} $\goodWSig{\Gm}{\wsig}{n}$ 
composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}).
%\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types". Since there are already boolean types and Σ types, Fin types can be defined. Using Fin, ordinary W types suffice to define trees with n constructors. So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1. We have to have enhanced W-type at FMLTT, that's for sure. Otherwise we (a). are too far away from the surface syntax because W-type itself only has one constructor all the time (b) this suggested encoding or syntactic sugar, since it interact with other connective like linkage (the Wrec is the interaction), so Wrec part is mostly unchanged  (c) since they will look the same, the encoding itself will look like an extra equality judgement in the syntax... which is not helpful for the complexity at all ... and even more complex  \\
%2. so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive. But my point is, it is hard to really write out the above syntactic sugar, formally. I think the issue here is similar to other reviewer saying "syntax translation". It is very hard to write out the encoding of Fin already, and then encode it again into W-type. (Just ask my agda mechanization). Of course, some hand-waiving can be done. For example, the part Zhixuan says I think it okay to hand-wave. But the destruction part -- the part Wrec interact with this encoding, I don't believe handwaving is enough. \\
%3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
%
Each pair models a constructor of the inductive type.
The $i$-th pair, projected from the signature~$\wsig$ using the forms
$\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
$\goodTerm{\Gm}{\wsigproj{i}{2}{\wsig}}{\wsigproj{i}{1}{\wsig} \to \cU}$,
defines the $i$-th constructor.
That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
$\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
$\wcode{\wsig}$ gives the code of the
W-type (\ruleref{tm/w}).

>>>>

Our formulation of W-types differs from previous ones in that it is
straightforward to identify constructors from what we call \emph{W-type signatures}.
%
A signature $\goodWSig{\Gm}{\wsig}{n}$ is
composed of $n$ pairs of types (\ruleref{wsig/empty} and \ruleref{wsig/add}),
%\Zhixuan{Page 14: I don't see the necessity of introducing "enhanced W-types". Since there are already boolean types and Σ types, Fin types can be defined. Using Fin, ordinary W types suffice to define trees with n constructors. So it might be clearer to just define 'enhanced W-types' as syntactic sugar.}\YZreply{I concur.}\EDJreply{This following sentence has been replied to Zhixuan and he doesn't has further response: 1. We have to have enhanced W-type at FMLTT, that's for sure. Otherwise we (a). are too far away from the surface syntax because W-type itself only has one constructor all the time (b) this suggested encoding or syntactic sugar, since it interact with other connective like linkage (the Wrec is the interaction), so Wrec part is mostly unchanged  (c) since they will look the same, the encoding itself will look like an extra equality judgement in the syntax... which is not helpful for the complexity at all ... and even more complex  \\
%2. so to me the problem you guys really have is around the content in the appendix -- I think Zhixuan mean he doesn't think this is more expressive. But my point is, it is hard to really write out the above syntactic sugar, formally. I think the issue here is similar to other reviewer saying "syntax translation". It is very hard to write out the encoding of Fin already, and then encode it again into W-type. (Just ask my agda mechanization). Of course, some hand-waiving can be done. For example, the part Zhixuan says I think it okay to hand-wave. But the destruction part -- the part Wrec interact with this encoding, I don't believe handwaving is enough. \\
%3. But how do I say this to the audience? I mean the main reason I didn't choose syntactic translation/syntax sugar is basically ``doing encoding is harder than directly working in ambient logic.'' }
%
each modeling a constructor of the inductive type.
The $i$-th pair, projected from the signature~$\wsig$ using the forms
$\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and
$\goodTerm{\Gm}{\wsigproj{i}{2}{\wsig}}{\wsigproj{i}{1}{\wsig} \to \cU}$,
defines the $i$-th constructor.
That is, given two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
$\goodTerm{\Gm, x : \El{\wsigproj{i}{2}{\wsig}(t_1)} }{t_2}{\El{\wcode{\wsig}}}$,
we can construct a term $\Wsup{i}{\wsig}{t_1}{x.t_2}$ of the W-type (\ruleref{tm/wsup}).
$\wcode{\wsig}$ gives the code of the
W-type (\ruleref{tm/w}).
